---
title: "Biodiversity simulation"
author: "Erika"
date: "12/2/2020"
output: html_document
---

## Introduction
This document includes the code to run the diversity simulations I use in the first conservation biology lab.

## Set things up
```{r}
rm(list = ls())
library(dplyr)
library(tidyr)
library(ggplot2)
library(vegan)
library(boot)
```

## Part 1. Set up simulation data

### Assumptions		
Our aim is to compare estimates of biodiversity from several indices with actual levels of diversity in differently structured commmunities in order to learn more about how the indices themselves work. 

Users will be able to set the actual number of species in the community, between 10 and 100 (by tens) = `NumSpecies`
Users will be able to choose the % of the community sampled in the "field" = `PercentSampled`


###Create communities
Start by creating the communities. Each of our communities will consist of 10,000 individuals

Let's create some species that we can use to build communities

```{r}
NumSpecies<-10
SpeciesCode<-paste("Species_", 1:NumSpecies, sep = "")
```

### Build Communities
Community 1 is going to be one with an even distribution of species.  Then we will randomly select from that community.

So let's take our Species varaible  and make a vector that repeats each species to approximately 10,000 total.

```{r}
myMultiplier<-10000/NumSpecies
NumIndividuals<-10000
Community_1<-rep(SpeciesCode, each = myMultiplier)
```
The second community will be dominated by one species; there will a predominace of  individuals of one species just a few of each of the other species.
```{r}
#first, figure out the balance which depends on the number of species NumSpecies
Balance<-NumIndividuals/NumSpecies
Common<-NumIndividuals*(1-(1/NumSpecies))
Remaining<-NumIndividuals-Common
FreqEachRemainSp<-trunc(Remaining/(NumSpecies-1))
TotalRare<-FreqEachRemainSp*(NumSpecies -1)
Remainder<-Remaining - TotalRare
Break1<-NumIndividuals-((FreqEachRemainSp*(NumSpecies-2)))

#Now build community using those variables
Community_2<-vector(mode = "character", length = 10000)
Spec1<-rep(SpeciesCode[1], each = Common)
Community_2[1:Common]<-Spec1
if(Remainder == 0){
	OtherSpec<-rep(SpeciesCode[2:NumSpecies], each = FreqEachRemainSp)
	Community_2[Common+1:NumIndividuals]<-OtherSpec
} else {
	OtherSpec<-rep(SpeciesCode[2:(NumSpecies-1)], each = FreqEachRemainSp)
	LastSpec<-rep(SpeciesCode[NumSpecies], each = FreqEachRemainSp+Remainder)
	Community_2[(Common+1):(Common+(length(OtherSpec)))]<-OtherSpec
	Community_2[((Common+(length(OtherSpec)))+1):NumIndividuals]<-LastSpec
}

```
The third community will be one that has 10,000 individuals selected at random from among the species.

```{r}
Community_3<-sample(SpeciesCode, size = 10000, replace = T)
```

Now put them together into a data frame
```{r}
Communities<-data.frame(cbind(Community_1, Community_2, Community_3))
```

## Part 2. Simulations!

###Simulation 1. Species Richness Index
Let's calculate species richness from each community a bunch of times and compare average to actual.


Confirms that the actual species richness in each community is 100 species.

How well does the SR measure capture actual SR depends on how many samples are taken.

Generate 10 sample sizes that the user will choose from

```{r}
SS<-c(10, 100, 1000, 2000, 4000, 6000, 8000)
PerPop<-SS/10000
```
Now, for each sample size, randomly select that number of individuals from the community and calculate SR. Do this a lot of times to generate mean and range.

Set up vectors to accept Species Richness for each community
```{r}
Richness1<-vector(mode = "integer", length = length(SS))
Richness2<-vector(mode = "integer", length = length(SS))
Richness3<-vector(mode = "integer", length = length(SS))
```
Now set up a dataframe for each community to store results of a whole bunch of iterations.

```{r}
NumIterations<-10
RowNames<-paste("Iteration_", 1:NumIterations, sep = "")
ColNames<-SS

 for (i in 1:3){
 	assign(paste0("Rvals_Comm", i), data.frame(matrix(nrow = NumIterations, ncol = length(SS)), row.names = RowNames))
 	
 }
```


Now create the loop to do the sampling

Need an outer loop that steps through the iterations and an inner loop that steps through 1 iteration for each sample size

```{r}
#outer loop
for(i in 1:NumIterations){
	for (s in 1:length(SS)){
	Take1<-sample(Communities$Community_1, size=SS[s], replace = F)
	Take2<-sample(Communities$Community_2, size = SS[s], replace = F)
	Take3<-sample(Communities$Community_3, size = SS[s], replace = F)

Richness1[s]<-length(unique(Take1))
Richness2[s]<-length(unique(Take2))
Richness3[s]<-length(unique(Take3))
	}
	Rvals_Comm1[i,]<-Richness1
	Rvals_Comm2[i,]<-Richness2
	Rvals_Comm3[i,]<-Richness3
}
names(Rvals_Comm1)<-ColNames
names(Rvals_Comm2)<-ColNames
names(Rvals_Comm3)<-ColNames
```
Reshape data frames using tidyr
```{r}
 #add column for iteration number to make tidying easier
 Rvals_Comm1$Iteration<-1:NumIterations
 Rvals_Comm2$Iteration<-1:NumIterations
 Rvals_Comm3$Iteration<-1:NumIterations
 
 Com1_SR<-Rvals_Comm1 %>% pivot_longer(!Iteration, names_to = "SS", values_to = "SR")
 Com2_SR<-Rvals_Comm2 %>% pivot_longer(!Iteration, names_to = "SS", values_to = "SR")
 Com3_SR<-Rvals_Comm3 %>% pivot_longer(!Iteration, names_to = "SS", values_to = "SR")
 
 #now add community
 Com1_SR$Community<-1
 Com2_SR$Community<-2
 Com3_SR$Community<-3
 #now for tidyness, put community as left-most column via dplyr
 Com1_SR<-relocate(Com1_SR, Community)
 Com2_SR<-relocate(Com2_SR, Community)
 Com3_SR<-relocate(Com3_SR, Community)
```
Now join all three data frames

```{r}
All<-bind_rows(Com1_SR, Com2_SR, Com3_SR)
```
**This all works but is very ugly code, to be cleaned up at a future date when I have time.**


Now plot
```{r}
#df to hold actual SR
Actual<-data.frame(
	Community = factor(length(All$Community)),
	SS = character(length(All$Community)), 
	SR = integer(length(All$Community))
	)

Actual$Community<-as.factor(rep(1:3, each = (length(All$Community)/3)))
Actual$SS<-rep(SS, times = 3)
Actual$Actual_SR<-NumSpecies
Actual$Actual_SR<-as.integer(Actual$Actual_SR)

All$Community<-as.factor(All$Community)
All$SS<-as.numeric(All$SS)
RichnessPlot<-ggplot(data = NULL, aes(x = SS, y = SR, color = Community))+
	geom_point(All, mapping = aes(group =Iteration))+
	geom_jitter(All, mapping = aes(group = Iteration), width = 0.25)+
	geom_smooth(data = All, method = NULL)+
	geom_point(Actual, mapping= aes(x = SS, y = Actual_SR), color = "black")+
	scale_x_log10(breaks = SS)


RichnessPlot
```



