---
title: "Biodiversity"
author: "Erika"
date: "1/14/2019"
output: html_document
---

##Introduction

In conservation biology during the first lab, we calculate simpson's and shannon-weiner indices for a series of simulations and then as homework with some real data.

I'd like to step it up a bit and use R for some of the diversity calculations.  Doing so will require the vegan package.

###Step 1.  Set up your workspace
```{r}
rm(list = ls())
library(dplyr)
library(vegan)
library(ggplot2)
```

###Step 2. Pull in some data

```{r}
library(readxl)
DF <- read_excel("tree_data.xlsx")
```
### Step 3. Check data for issues

```{r}
DF$Species<-as.factor(DF$Species)
levels(DF$Species)
```
Looks good - each species just has one "name". No typos.

We don't really need the DBH data for this one.

###Step 4. Determine # of each species
```{r}
Species <- DF %>% group_by(Species) %>% summarize (
	NumSpec=length(Species)
)
```
###Step 5. Transpose data to vegan format
Need to have columns represent species and row represent # per species

```{r}
DF1<-t(Species)
DF1<-as.data.frame(DF1)
NumSpec<-DF1[2,]
NumSpec<-as.numeric(NumSpec)
```
###Step 6. Calculate diversity indices
```{r}
diversity(NumSpec, index = "simpson")
```


